---
title: "TPM calculation"
author: "Laura Toribio-Celestino"
date: "10/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading required packages

```{r, results='hide', eval=TRUE, message=FALSE, warning=FALSE}
library(DESeq2)
library(dplyr)
library(ggplot2)
library(reshape2)
library(dunn.test)
```

### Importing the count table and constructing the DESeq2 object

```{r, eval=TRUE}
# Importing count table
fc_full <- read.table("featureCounts.tsv", header = TRUE, row.names = 1)
colnames(fc_full) <- gsub("\\.bam$", "", colnames(fc_full))
countdata <- fc_full
# Saving gene lengths and chromosome ids
geneLength <- countdata$Length
chrName <- countdata$Chr
# Final count tables
countdata$Length <- NULL
countdata$Chr <- NULL
head(countdata)
tail(countdata)
```

Building the variables for the columns of the colData object.

```{r, eval=TRUE}
condition <- c('wt', 'wt', 'wt', 'cur', 'cur', 'cur')
replicate <- c('1', '2', '3', '1', '2', '3')
coldata <- as.data.frame(cbind(colnames(countdata), condition, replicate))
coldata
```

Constructing the DESeqDataSet object. Samples that do not carry pOXA-48 are removed after constructing the dds objects.

```{r, eval=TRUE}
plasfreenames <- c("H53c1.1", "H53c1.2", "H53c1.3")
dds <- DESeqDataSetFromMatrix(countData = countdata, colData = coldata, design = ~ condition)
dds <- dds[, !colnames(dds) %in% plasfreenames]
dds
```

### Calculating FPKM values

```{r, eval=TRUE}
mcols(dds)$basepairs = geneLength
fpkm_all <- fpkm(dds)
head(fpkm_all)
tail(fpkm_all)
```

### Calculating TPM values

```{r, eval=TRUE}
tpm_all <- as.data.frame(apply(fpkm_all, 2, function(x){(x/sum(x))*1E6}))
tpm_all$Chr <- chrName
head(tpm_all)
tail(tpm_all)
```

### Violin plots

```{r, eval=TRUE}
df <- tibble::rownames_to_column(tpm_all, "GeneID")
df[df == "cds-pgaptmp_005364"] <- "blaOXA-48"

# Only chromosomal and pOXA-48 rows are selected
df <- subset(df, Chr %in% c("1", "1;1", "3"))
df$Chr[df$Chr %in% c("1", "1;1")] <- "Chromosome"
df$Chr[df$Chr %in% c("3")] <- "pOXA-48"
# Dataframe is melted for correct representation
dfmelted <- melt(df, id.vars=c('Chr', 'GeneID'), value.name = 'TPM', variable.name = 'Replicate')
head(dfmelted)
tail(dfmelted)

p <- ggplot(dfmelted, aes(Replicate, TPM))
p + geom_violin(aes(Replicate, TPM, fill=Chr), position=position_dodge(.9)) +
  xlab(NULL) + ylab("log2 TPM") + scale_y_continuous(trans = "log10") +
  stat_summary(aes(group=Chr), position=position_dodge(.9), fun = "median", geom = "point", color = "black") +
  geom_text(aes(Replicate, TPM, label=GeneID), data=dfmelted[dfmelted$GeneID %in% "blaOXA-48", ])
```

### Some stats

Spearman correlation between pOXA-48-carrying replicates.

```{r, eval=TRUE}
cor(tpm_all[,1:3], method="spearman")
```

Kruskal-Wallis test followed by Dunn's test to assess differences between replicates.

```{r, eval=TRUE}
kruskal.test(TPM ~ Replicate, data = dfmelted)
dunn.test(dfmelted$TPM, g = dfmelted$Replicate, method = "bh")
```

### Calculating mean TPM values

```{r, eval=TRUE}
df$median_tpm <- apply(df[, 2:4], 1, median) # change replicate columns if necessary
head(df)
tail(df)
write.table(df, file="TPM_H53_chr_pOXA48.tsv", sep="\t", row.names=FALSE)
```

# Session Info

```{r, eval=TRUE}
sessionInfo()
```